Nonlinear Spatio-Temporal Modelling using the Integro-Difference Equation

### 2010-07-25 ltx/second_order_framework/second_order_manuscript.tex
correction to EM algorithm

\usepackage{anysize} 									%<-use this for checking equation length
\marginsize {1.7cm}{1.5cm}{1cm}{1cm} 				%<-use this for checking equation length

@@

The $p^{th}$-order nonlinear, stochastic, spatio-temporal homogenous IDE model is defined as

@@

  \rho_t(\mathbf r)=\int_{\Omega}w(\mathbf r-\mathbf r')f(v_{t-p}(\mathbf r'))d\mathbf r'+\mathbf e_{t-p}(\mathbf r). 

@@

where $v_t(\mathbf r)$ denotes a spatio-temporal field at time $t$ and at spatial location $\mathbf r$, $f(\cdot)$ is a nonlinear map, and $w(\cdot)$ is a spatial mixing kernel. The disturbance, $\mathbf e_{t-p}(\mathbf r)$, is a zero-mean normally distributed noise process that is spatially

@@

Observations from the spatio-temporal field are described by the measurement function

@@

where the observation vector, $\mathbf{y}_{t}$, compiled at $n_{y}$ spatial locations and at time $t$, is corrupted by an i.i.d normally distributed zero-mean white noise ${\boldsymbol\varepsilon}_t\sim \mathcal{N}\left(\mathbf{0},\boldsymbol\Sigma_{\varepsilon}\right)$ with $\mathbf{\Sigma}_{\boldsymbol\varepsilon}=\sigma_{\varepsilon}^2\mathbf I_{n_y} $. It is assumed here that $\boldsymbol\Sigma_{\varepsilon}$ and $\gamma\left(\mathbf r-\mathbf r' \right)$ are known. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the sensor pick-up geometry and is defined by 

@@

	m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)},

@@

where $\sigma_m$ sets the sensor width. The superscript $\top$ denotes the transpose operator. We proceed by writing the spatio-temporal IDE model in a parameterized state space form which allows the application of standard estimation techniques. We use a decomposition of the field and the spatial mixing kernel using a set of Gaussian basis functions

@@

	\label{eq:FieldDecomp} v_t\left(\mathbf{r}\right) \approx \boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t

@@

	 w\left(\mathbf{r}-\mathbf{r}'\right) \approx\boldsymbol{\psi}^\top\left(\mathbf{r}-\mathbf{r}'\right) \boldsymbol{\theta}

@@

where $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$ and $\mathbf{\boldsymbol{\psi}}(\mathbf{r})$ are  vectors of Gaussian basis functions that are scaled by the state vector, $\mathbf{x}_t$ and the scaling parameter vector $\boldsymbol\theta$ respectively. Substituting \eqref{eq:LagPolynomial} in \eqref{eq:IDEModel} and using the above decomposition we have

@@

\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right)(1+\sum_{i=1}^{p}\varphi_i L_i) \mathbf{x}_t=\int_{\Omega}\boldsymbol{\psi}^\top\left(\mathbf{r}-\mathbf{r}'\right) f(\boldsymbol{\phi}^{\top}\left(\mathbf{r'}\right) \mathbf{x}_{t-p})d\mathbf r'\boldsymbol{\theta}+\mathbf e_{t-p}(\mathbf r)

@@

multiplying \eqref{eq:DecompModel} by $\boldsymbol \phi(\mathbf r)$ and integrate over the spatial domain $\Omega$ we get

@@

 \boldsymbol \Gamma (1+\sum_{i=1}^{p}\varphi_i L_i) \mathbf{x}_t=\int_{\Omega}\boldsymbol \phi(\mathbf r)\int_{\Omega}\boldsymbol{\psi}^\top\left(\mathbf{r}-\mathbf{r}'\right) f(\boldsymbol{\phi}^{\top}\left(\mathbf{r'}\right) \mathbf{x}_{t-p})d\mathbf r'\boldsymbol{\theta}+\int_{\Omega}\boldsymbol\phi(\mathbf r)\mathbf e_{t-p}(\mathbf r)d\mathbf r

@@

	\boldsymbol{\Gamma} \triangleq \int_\Omega {\boldsymbol{\phi} \left(\mathbf{r}\right)\boldsymbol{\phi} ^{\top}\left(\mathbf{r}\right)d\mathbf{r}} 

@@

cross-multiplying \eqref{eq:GammaDecompModel} by $\boldsymbol{\Gamma}^{-1}$ gives

@@

 (1+\sum_{i=1}^{p}\varphi_i L_i) \mathbf{x}_t=\boldsymbol\Gamma^{-1}\int_{\Omega}\boldsymbol \phi(\mathbf r)\int_{\Omega}\boldsymbol{\psi}^\top\left(\mathbf{r}-\mathbf{r}'\right) f(\boldsymbol{\phi}^{\top}\left(\mathbf{r'}\right) \mathbf{x}_{t-p})d\mathbf r'\boldsymbol{\theta}+\boldsymbol\Gamma^{-1}\int_{\Omega}\boldsymbol\phi(\mathbf r)\mathbf e_{t-p}(\mathbf r)d\mathbf r

@@

This can be further simplified by assuming isotropic structure for the spatial kernel

@@

	\boldsymbol{\psi} (\mathbf{r}-\mathbf{r}') = \boldsymbol{\psi} (2c+\mathbf{r}'-\mathbf{r}).

@@

where $c$ is the center of the kernel. \dean{Can't use c as it is used in the observation matrix.} To make the simplification, we first define

@@

	\boldsymbol{\Psi}(\mathbf{r}') \triangleq T_s\boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\psi}^{\top} (2c+\mathbf{r}'-\mathbf{r})\textrm{d}\mathbf{r}},

@@

which is a constant $n_x \times n_{\theta}$ matrix and can be computed analytically. Now substituting \eqref{eq:DefPsi} into \eqref{eq:InvGammaDecompModel} we get

@@

(1+\sum_{i=1}^{p}\varphi_i L_i) \mathbf{x}_t = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \boldsymbol{\theta}
+ \boldsymbol{\Gamma}^{-1} \int_\Omega{\boldsymbol{\phi}(\mathbf{r})e_{t-p}(\mathbf{r})\textrm{d}\mathbf{r}}.

@@

	\mathbf{e}_{t-p} \triangleq \boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi} ( \mathbf{r} )e_{t-p}( \mathbf{r} )\textrm{d}\mathbf{r}},

@@

	\boldsymbol\Sigma_e =\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \gamma\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top}. 

@@

the observation equation of the reduced model is found by substituting \eqref{eq:FieldDecomp} into \eqref{eq:ObservationEquation} giving

@@

	c_{i,j} = \int_{\Omega}m(\mathbf{r}_i - \mathbf{r}')\boldsymbol{\phi}_j(\mathbf{r}')\textrm{d}\mathbf{r}'. 

@@

	(1+\sum_{i=1}^{p}\varphi_i L_i) \mathbf{x}_t = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_{t-p}) \textrm{d}\mathbf{r}' \boldsymbol{\theta}

@@

x_t+\varphi_{1} x_{t-1}=\boldsymbol\Gamma^{-1}\int_{\Omega}\boldsymbol \phi(\mathbf r)\int_{\Omega}\boldsymbol{\psi}^\top\left(\mathbf{r}-\mathbf{r}'\right) f(\boldsymbol{\phi}^{\top}\left(\mathbf{r'}\right) \mathbf{x}_{t-1})d\mathbf r'\boldsymbol{\theta}+\boldsymbol\Gamma^{-1}\int_{\Omega}\boldsymbol\phi(\mathbf r)\mathbf e_{t-1}(\mathbf r)d\mathbf r

@@

	v\left( {\mathbf{r},t} \right) = \int_{ - \infty }^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'}

@@

where $\zeta=\tau^{-1}$ and $\tau$ is the synaptic time constant and $\eta(t)$ is the Heaviside step function. $A$ determines the maximal amplitude of the post-synaptic potentials.

@@

	g\left( \mathbf{r},t \right) = \int_\Omega  {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right) \right)d\mathbf{r}'} 

@@

where $f(.)$ is the firing rate function, $w(.)$ is the spatial connectivity kernel, and $\Omega$ is the spatial domain.

@@

 \frac{dh(t)}{dt}&=&A[\delta(t)t\mathrm{exp}(-\zeta t)+\eta(t)\frac{d}{dt}(t\mathrm{exp}(-\zeta t))] \nonumber\\
&=& A[0+\eta(t)[\mathrm{exp}(-\zeta t)-\zeta t\mathrm{exp}(-\zeta t) ]\nonumber \\

@@

 \frac{d^2h(t)}{dt^2}&=&A[\delta(t)\mathrm{exp}(-\zeta t)-\zeta\eta(t)\mathrm{exp}(-\zeta t)]-\zeta \frac{dh(t)}{dt} \nonumber\\
&=& A\delta(t)-\zeta A \eta(t)\mathrm{exp}(-\zeta t)-\zeta \frac{dh(t)}{dt} \nonumber \\

@@

We define the operator $D$ as $D=\frac{d^2}{dt^2}+2\zeta \frac{d}{dt}+\zeta^2 $, applying $D$ on $h$ and using the result from (\ref{SynapticRespKernelSecondderivation}) it is clear that 

@@

 \frac{d^2v(\mathbf r,t)}{dt^2}+2\zeta \frac{dv(\mathbf r,t)}{dt}+\zeta^2v(\mathbf r,t)&=Ag\left( \mathbf{r},t \right)\\
&=A\int_\Omega  {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right) \right)d\mathbf{r}'} \\

@@

The above second-order differential equation can be split into two, coupled first order ODEs.\\

@@

\tilde{v}_{t+1}=(1-2\zeta T_s)\tilde{v}_t-\zeta^2T_sv_t+AT_sg_t\left( \mathbf{r}\right)

@@

 \begin{bmatrix}v_{t+1} \\ \tilde{v}_{t+1}\end{bmatrix}= \begin{bmatrix} T_s &1\\ 1-2\zeta T_s & -\zeta^2T_s\end{bmatrix}\begin{bmatrix}v_{t} \\ \tilde{v}_{t}\end{bmatrix}+AT_sg_t\left( \mathbf{r}\right)

@@

 v_{t+2}-2(1-\zeta T_s)v_{t+1}+(1-2\zeta T_s+\zeta^2T_s^2)v_t=AT_s^2g_t\left( \mathbf{r}\right)

@@

 v_{t+2}-2v_{t+1}+2\zeta T_S v_{t+1}+v_t-2\zeta T_sv_t+\zeta^2T_s^2v_t=AT_s^2g_t\left( \mathbf{r}\right)

@@

(v_{t+2}-v_{t+1})-(v_{t+1}-v_t)+2\zeta T_s(v_{t+1}-v_t)+\zeta^2T_s^2v_t=AT_s^2g_t\left( \mathbf{r}\right)

@@

(v_{t+2}-v_{t+1})-(1-2\zeta T_s)(v_{t+1}-v_t)+\zeta^2T_s^2v_t=AT_s^2g_t\left( \mathbf{r}\right)

@@

substituting \eqref{eq:tildev} in the above equation and multiplying by $1/T_s$ we get the state space model

@@

The model relates the average number of action potentials $g(\mathbf{r},t)$ arriving at time $t \in \mathbb R^{+}$ and at position $\mathbf{r}$ to the local post-synaptic membrane voltage $v(\mathbf{r},t)$. The post-synaptic potentials generated at a neuronal population at location $\mathbf{r}$ by action potentials arriving from all other connected populations at locations $\mathbf{r}'$ is described by 

@@

	\label{eq:SpikesToPotential} v\left( {\mathbf{r},t} \right)=v(\mathbf r, 0)h(t)+\int_0^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'} 

@@

Let $W(\mathbf r,t)$ be a spatio-temporal Wiener process. Then, for $\sigma \in \mathbb R^{+}$, the non-local interactions $g\left( {\mathbf{r},t} \right)$ between cortical populations can be divided into a predominantly deterministic part $g'\left( {\mathbf{r},t} \right) $ and  the random component $W(\mathbf r,t)$,

@@

  g\left( {\mathbf{r},t} \right)dt=g'\left( {\mathbf{r},t} \right)dt+\sigma d W(r,t)

@@

As a result, the post-synaptic membrane voltage $v(\mathbf r, t)$ is itself a stochastic process. The deterministic part $g'(\mathbf r, t)$ is usually modelled as an integral equation with exogenous deterministic input , $ u(\mathbf r,t)$, of the form \cite{Atay2005}

@@

	\label{DeterministicRateBasedInteractions} g'\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right) \right)\textrm{d}\mathbf{r}'}+u(\mathbf r,t)

@@

where $f(\cdot)$ is the firing rate function, $w(\cdot)$ is the spatial connectivity kernel and $\Omega$ is the spatial domain representing a cortical sheet or surface. Substituting \eqref{eq:ActPotComponents} in \eqref{eq:SpikesToPotential} we get

@@

v\left( {\mathbf{r},t} \right)=v_0h(t)+\int_0^t {h\left( {t - t'} \right)g'\left( {\mathbf{r},t'} \right)dt'}+ \int_0^t h\left( {t - t'} \right)\sigma d W(r,t')

@@

Here we assume that $v_0$ is generated by a known distribution $p_0(v)$ and is independent of the the random component $W(\mathbf r,t)$. We stress that $\sigma$ does not depend on the field $v(\mathbf r, t)$
and hence the noise in \eqref{eq:NeuralModelDeterStoch} is strictly additive. It can be shown that the post-synaptic potential $v(\mathbf r, t)$ solves the following initial value problem (see Appendix \ref{app:InitialValueProblemProof})

@@

 dv(\mathbf r, t) + \zeta v(\mathbf r, t)dt = g'(\mathbf r, t)dt + dW(\mathbf r, t) \quad t\ge0, v(\mathbf r, 0) = v_0

@@

applying the Euler-Maruyama method the integro-difference equation (IDE) model takes the form of

@@

 v_{t+T_s}(\mathbf r)- v_{t}(\mathbf r)+ T_s\zeta v_t(\mathbf r)= T_sg'(\mathbf r, t) + (W_{t+T_s}(\mathbf r)-W_t(\mathbf r))

@@

where $T_s$ is the time step, the increments $W_{t+T_s}(\mathbf r)-W_t(\mathbf r)$ is normally distributed with mean zero and covariance function $T_s\gamma(\mathbf r-\mathbf r')$. We rewrite \eqref{eq:Euler-MaruyamaDiscrit} as

@@

 where $\xi = 1-\zeta T_s$ and $e_t(\mathbf{r})$ is an $i.i.d.$ disturbance such that $e_t(\mathbf{r})\sim\mathcal{GP}(\mathbf 0,T_s\gamma(\mathbf{r}-\mathbf{r}'))$. Here $\mathcal{GP}(\mathbf 0,T_s\gamma(\mathbf{r}-\mathbf{r}'))$ denotes a zero mean spatial Gaussian process with covariance function $T_s\gamma(\mathbf{r}-\mathbf{r}')$. To simplify the notation, the index of the future time sample, $t+T_s$, shall be referred to as $t+1$ throughout the rest of the paper. 

@@

	\label{eq:SecondOrderSpikesToPotential} v\left( {\mathbf{r},t} \right)=v(\mathbf r, 0)(t+1)e^{-\zeta t}+v'(\mathbf r,0)te^{-\zeta t}+\int_0^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'} 

@@

	\label{eq:SecondOrderSynapticRespKernel} h(t) = At\exp{\left(-\zeta t\right)} 

@@

where $\zeta=\tau^{-1}$, $\tau$ is the synaptic time constant.  $A$ determines the maximal amplitude of the post-synaptic potentials.
Let $W(\mathbf r,t)$ be a spatio-temporal Wiener process. Then, for $\sigma \in \mathbb R^{+}$, the non-local interactions $g\left( {\mathbf{r},t} \right)$ between cortical populations can be divided into a predominantly deterministic part $g'\left( {\mathbf{r},t} \right) $ and  the random component $W(\mathbf r,t)$,

@@

where $f(\cdot)$ is the firing rate function, $w(\cdot)$ is the spatial connectivity kernel and $\Omega$ is the spatial domain representing a cortical sheet or surface. Substituting \eqref{eq:SecondOrderActPotComponents} in \eqref{eq:SecondOrderSpikesToPotential} we get

@@

It can be shown that the post-synaptic potential $v(\mathbf r, t)$ solves the following initial value problem \parham{Andrew do you know how?}

>>>>

\usepackage{anysize} 								
%<-use this for checking equation length
\marginsize {1.7cm}{1.5cm}{1cm}{1cm} 				%<-use this for
% checking equation length

@@

The $p^{th}$-order nonlinear, stochastic, spatio-temporal homogenous IDE model
is defined as

@@

  \rho_t(\mathbf r)=\int_{\Omega}w(\mathbf r-\mathbf r')f(v_{t-p}(\mathbf
r'))d\mathbf r'+\mathbf e_{t-p}(\mathbf r). 

@@

where $v_t(\mathbf r)$ denotes a spatio-temporal field at time $t$ and at
spatial location $\mathbf r$, $f(\cdot)$ is a nonlinear map, and $w(\cdot)$ is a
spatial mixing kernel. The disturbance, $\mathbf e_{t-p}(\mathbf r)$, is a
zero-mean normally distributed noise process that is spatially

@@

Observations from the spatio-temporal field are described by the measurement
function

@@

where the observation vector, $\mathbf{y}_{t}$, compiled at $n_{y}$ spatial
locations and at time $t$, is corrupted by an i.i.d normally distributed
zero-mean white noise ${\boldsymbol\varepsilon}_t\sim
\mathcal{N}\left(\mathbf{0},\boldsymbol\Sigma_{\varepsilon}\right)$ with
$\mathbf{\Sigma}_{\boldsymbol\varepsilon}=\sigma_{\varepsilon}^2\mathbf I_{n_y}
$. It is assumed here that $\boldsymbol\Sigma_{\varepsilon}$ and
$\gamma\left(\mathbf r-\mathbf r' \right)$ are known. The output kernel
$m(\mathbf{r}-\mathbf{r}')$ governs the sensor pick-up geometry and is defined
by 

@@

	m\left(\mathbf{r}-\mathbf{r}'\right) =
\exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{
\sigma_m^2}\right)},

@@

where $\sigma_m$ sets the sensor width. The superscript $\top$ denotes the
transpose operator. We proceed by writing the spatio-temporal IDE model in a
parameterized state space form which allows the application of standard
estimation techniques. We use a decomposition of the field and the spatial
mixing kernel using a set of Gaussian basis functions

@@

	\label{eq:FieldDecomp} v_t\left(\mathbf{r}\right) \approx
\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t

@@

	 w\left(\mathbf{r}-\mathbf{r}'\right)
\approx\boldsymbol{\psi}^\top\left(\mathbf{r}-\mathbf{r}'\right)
\boldsymbol{\theta}

@@

where $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$ and
$\mathbf{\boldsymbol{\psi}}(\mathbf{r})$ are  vectors of Gaussian basis
functions that are scaled by the state vector, $\mathbf{x}_t$ and the scaling
parameter vector $\boldsymbol\theta$ respectively. Substituting
\eqref{eq:LagPolynomial} in \eqref{eq:IDEModel} and using the above
decomposition we have

@@

\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right)(1+\sum_{i=1}^{p}\varphi_i L_i)
\mathbf{x}_t=\int_{\Omega}\boldsymbol{\psi}^\top\left(\mathbf{r}-\mathbf{r}
'\right) f(\boldsymbol{\phi}^{\top}\left(\mathbf{r'}\right)
\mathbf{x}_{t-p})d\mathbf r'\boldsymbol{\theta}+\mathbf e_{t-p}(\mathbf r)

@@

multiplying \eqref{eq:DecompModel} by $\boldsymbol \phi(\mathbf r)$ and
integrate over the spatial domain $\Omega$ we get

@@

 \boldsymbol \Gamma (1+\sum_{i=1}^{p}\varphi_i L_i)
\mathbf{x}_t=\int_{\Omega}\boldsymbol \phi(\mathbf
r)\int_{\Omega}\boldsymbol{\psi}^\top\left(\mathbf{r}-\mathbf{r}'\right)
f(\boldsymbol{\phi}^{\top}\left(\mathbf{r'}\right) \mathbf{x}_{t-p})d\mathbf
r'\boldsymbol{\theta}+\int_{\Omega}\boldsymbol\phi(\mathbf r)\mathbf
e_{t-p}(\mathbf r)d\mathbf r

@@

	\boldsymbol{\Gamma} \triangleq \int_\Omega {\boldsymbol{\phi}
\left(\mathbf{r}\right)\boldsymbol{\phi}
^{\top}\left(\mathbf{r}\right)d\mathbf{r}} 

@@

cross-multiplying \eqref{eq:GammaDecompModel} by $\boldsymbol{\Gamma}^{-1}$
gives

@@

 (1+\sum_{i=1}^{p}\varphi_i L_i)
\mathbf{x}_t=\boldsymbol\Gamma^{-1}\int_{\Omega}\boldsymbol \phi(\mathbf
r)\int_{\Omega}\boldsymbol{\psi}^\top\left(\mathbf{r}-\mathbf{r}'\right)
f(\boldsymbol{\phi}^{\top}\left(\mathbf{r'}\right) \mathbf{x}_{t-p})d\mathbf
r'\boldsymbol{\theta}+\boldsymbol\Gamma^{-1}\int_{\Omega}\boldsymbol\phi(\mathbf
r)\mathbf e_{t-p}(\mathbf r)d\mathbf r

@@

This can be further simplified by assuming isotropic structure for the spatial
kernel

@@

	\boldsymbol{\psi} (\mathbf{r}-\mathbf{r}') = \boldsymbol{\psi}
(2c+\mathbf{r}'-\mathbf{r}).

@@

where $c$ is the center of the kernel. \dean{Can't use c as it is used in the
observation matrix.} To make the simplification, we first define

@@

	\boldsymbol{\Psi}(\mathbf{r}') \triangleq
T_s\boldsymbol{\Gamma}^{-1}\int_\Omega
{\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\psi}^{\top}
(2c+\mathbf{r}'-\mathbf{r})\textrm{d}\mathbf{r}},

@@

which is a constant $n_x \times n_{\theta}$ matrix and can be computed
analytically. Now substituting \eqref{eq:DefPsi} into
\eqref{eq:InvGammaDecompModel} we get

@@

(1+\sum_{i=1}^{p}\varphi_i L_i) \mathbf{x}_t = \int_\Omega
\boldsymbol{\Psi}(\mathbf{r}')
f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}'
\boldsymbol{\theta}
+ \boldsymbol{\Gamma}^{-1}
\int_\Omega{\boldsymbol{\phi}(\mathbf{r})e_{t-p}(\mathbf{r})\textrm{d}\mathbf{r}
}.

@@

	\mathbf{e}_{t-p} \triangleq \boldsymbol{\Gamma}^{-1}\int_\Omega
{\boldsymbol{\phi} ( \mathbf{r} )e_{t-p}( \mathbf{r} )\textrm{d}\mathbf{r}},

@@

	\boldsymbol\Sigma_e
=\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf
r\right) \gamma\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf
r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top}. 

@@

the observation equation of the reduced model is found by substituting
\eqref{eq:FieldDecomp} into \eqref{eq:ObservationEquation} giving

@@

	c_{i,j} = \int_{\Omega}m(\mathbf{r}_i -
\mathbf{r}')\boldsymbol{\phi}_j(\mathbf{r}')\textrm{d}\mathbf{r}'. 

@@

	(1+\sum_{i=1}^{p}\varphi_i L_i) \mathbf{x}_t = \int_\Omega
\boldsymbol{\Psi}(\mathbf{r}')
f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_{t-p}) \textrm{d}\mathbf{r}'
\boldsymbol{\theta}

@@

x_t+\varphi_{1} x_{t-1}=\boldsymbol\Gamma^{-1}\int_{\Omega}\boldsymbol
\phi(\mathbf
r)\int_{\Omega}\boldsymbol{\psi}^\top\left(\mathbf{r}-\mathbf{r}'\right)
f(\boldsymbol{\phi}^{\top}\left(\mathbf{r'}\right) \mathbf{x}_{t-1})d\mathbf
r'\boldsymbol{\theta}+\boldsymbol\Gamma^{-1}\int_{\Omega}\boldsymbol\phi(\mathbf
r)\mathbf e_{t-1}(\mathbf r)d\mathbf r

@@

	v\left( {\mathbf{r},t} \right) = \int_{ - \infty }^t {h\left( {t - t'}
\right)g\left( {\mathbf{r},t'} \right)dt'}

@@

where $\zeta=\tau^{-1}$ and $\tau$ is the synaptic time constant and $\eta(t)$
is the Heaviside step function. $A$ determines the maximal amplitude of the
post-synaptic potentials.

@@

	g\left( \mathbf{r},t \right) = \int_\Omega  {w\left(
\mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right)
\right)d\mathbf{r}'} 

@@

where $f(.)$ is the firing rate function, $w(.)$ is the spatial connectivity
kernel, and $\Omega$ is the spatial domain.

@@

 \frac{dh(t)}{dt}&=&A[\delta(t)t\mathrm{exp}(-\zeta
t)+\eta(t)\frac{d}{dt}(t\mathrm{exp}(-\zeta t))] \nonumber\\
&=& A[0+\eta(t)[\mathrm{exp}(-\zeta t)-\zeta t\mathrm{exp}(-\zeta t) ]\nonumber
\\

@@

 \frac{d^2h(t)}{dt^2}&=&A[\delta(t)\mathrm{exp}(-\zeta
t)-\zeta\eta(t)\mathrm{exp}(-\zeta t)]-\zeta \frac{dh(t)}{dt} \nonumber\\
&=& A\delta(t)-\zeta A \eta(t)\mathrm{exp}(-\zeta t)-\zeta \frac{dh(t)}{dt}
\nonumber \\

@@

We define the operator $D$ as $D=\frac{d^2}{dt^2}+2\zeta \frac{d}{dt}+\zeta^2 $,
applying $D$ on $h$ and using the result from
(\ref{SynapticRespKernelSecondderivation}) it is clear that 

@@

 \frac{d^2v(\mathbf r,t)}{dt^2}+2\zeta \frac{dv(\mathbf
r,t)}{dt}+\zeta^2v(\mathbf r,t)&=Ag\left( \mathbf{r},t \right)\\
&=A\int_\Omega  {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left(
\mathbf{r}',t \right) \right)d\mathbf{r}'} \\

@@

The above second-order differential equation can be split into two, coupled
first order ODEs.\\

@@

\tilde{v}_{t+1}=(1-2\zeta T_s)\tilde{v}_t-\zeta^2T_sv_t+AT_sg_t\left(
\mathbf{r}\right)

@@

 \begin{bmatrix}v_{t+1} \\ \tilde{v}_{t+1}\end{bmatrix}= \begin{bmatrix} T_s
&1\\ 1-2\zeta T_s & -\zeta^2T_s\end{bmatrix}\begin{bmatrix}v_{t} \\
\tilde{v}_{t}\end{bmatrix}+AT_sg_t\left( \mathbf{r}\right)

@@

 v_{t+2}-2(1-\zeta T_s)v_{t+1}+(1-2\zeta T_s+\zeta^2T_s^2)v_t=AT_s^2g_t\left(
\mathbf{r}\right)

@@

 v_{t+2}-2v_{t+1}+2\zeta T_S v_{t+1}+v_t-2\zeta
T_sv_t+\zeta^2T_s^2v_t=AT_s^2g_t\left( \mathbf{r}\right)

@@

(v_{t+2}-v_{t+1})-(v_{t+1}-v_t)+2\zeta
T_s(v_{t+1}-v_t)+\zeta^2T_s^2v_t=AT_s^2g_t\left( \mathbf{r}\right)

@@

(v_{t+2}-v_{t+1})-(1-2\zeta T_s)(v_{t+1}-v_t)+\zeta^2T_s^2v_t=AT_s^2g_t\left(
\mathbf{r}\right)

@@

substituting \eqref{eq:tildev} in the above equation and multiplying by $1/T_s$
we get the state space model

@@

The model relates the average number of action potentials $g(\mathbf{r},t)$
arriving at time $t \in \mathbb R^{+}$ and at position $\mathbf{r}$ to the local
post-synaptic membrane voltage $v(\mathbf{r},t)$. The post-synaptic potentials
generated at a neuronal population at location $\mathbf{r}$ by action potentials
arriving from all other connected populations at locations $\mathbf{r}'$ is
described by 

@@

	\label{eq:SpikesToPotential} v\left( {\mathbf{r},t} \right)=v(\mathbf r,
0)h(t)+\int_0^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'} 

@@

Let $W(\mathbf r,t)$ be a spatio-temporal Wiener process. Then, for $\sigma \in
\mathbb R^{+}$, the non-local interactions $g\left( {\mathbf{r},t} \right)$
between cortical populations can be divided into a predominantly deterministic
part $g'\left( {\mathbf{r},t} \right) $ and  the random component $W(\mathbf
r,t)$,

@@

  g\left( {\mathbf{r},t} \right)dt=g'\left( {\mathbf{r},t} \right)dt+\sigma d
W(r,t)

@@

As a result, the post-synaptic membrane voltage $v(\mathbf r, t)$ is itself a
stochastic process. The deterministic part $g'(\mathbf r, t)$ is usually
modelled as an integral equation with exogenous deterministic input , $
u(\mathbf r,t)$, of the form \cite{Atay2005}

@@

	\label{DeterministicRateBasedInteractions} g'\left( \mathbf{r},t \right)
= \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left(
\mathbf{r}',t \right) \right)\textrm{d}\mathbf{r}'}+u(\mathbf r,t)

@@

where $f(\cdot)$ is the firing rate function, $w(\cdot)$ is the spatial
connectivity kernel and $\Omega$ is the spatial domain representing a cortical
sheet or surface. Substituting \eqref{eq:ActPotComponents} in
\eqref{eq:SpikesToPotential} we get

@@

v\left( {\mathbf{r},t} \right)=v_0h(t)+\int_0^t {h\left( {t - t'}
\right)g'\left( {\mathbf{r},t'} \right)dt'}+ \int_0^t h\left( {t - t'}
\right)\sigma d W(r,t')

@@

Here we assume that $v_0$ is generated by a known distribution $p_0(v)$ and is
independent of the the random component $W(\mathbf r,t)$. We stress that
$\sigma$ does not depend on the field $v(\mathbf r, t)$
and hence the noise in \eqref{eq:NeuralModelDeterStoch} is strictly additive. It
can be shown that the post-synaptic potential $v(\mathbf r, t)$ solves the
following initial value problem (see Appendix
\ref{app:InitialValueProblemProof})

@@

 dv(\mathbf r, t) + \zeta v(\mathbf r, t)dt = g'(\mathbf r, t)dt + dW(\mathbf r,
t) \quad t\ge0, v(\mathbf r, 0) = v_0

@@

applying the Euler-Maruyama method the integro-difference equation (IDE) model
takes the form of

@@

 v_{t+T_s}(\mathbf r)- v_{t}(\mathbf r)+ T_s\zeta v_t(\mathbf r)= T_sg'(\mathbf
r, t) + (W_{t+T_s}(\mathbf r)-W_t(\mathbf r))

@@

where $T_s$ is the time step, the increments $W_{t+T_s}(\mathbf r)-W_t(\mathbf
r)$ is normally distributed with mean zero and covariance function
$T_s\gamma(\mathbf r-\mathbf r')$. We rewrite \eqref{eq:Euler-MaruyamaDiscrit}
as

@@

 where $\xi = 1-\zeta T_s$ and $e_t(\mathbf{r})$ is an $i.i.d.$ disturbance such
that $e_t(\mathbf{r})\sim\mathcal{GP}(\mathbf
0,T_s\gamma(\mathbf{r}-\mathbf{r}'))$. Here $\mathcal{GP}(\mathbf
0,T_s\gamma(\mathbf{r}-\mathbf{r}'))$ denotes a zero mean spatial Gaussian
process with covariance function $T_s\gamma(\mathbf{r}-\mathbf{r}')$. To
simplify the notation, the index of the future time sample, $t+T_s$, shall be
referred to as $t+1$ throughout the rest of the paper. 

@@

	\label{eq:SecondOrderSpikesToPotential} v\left( {\mathbf{r},t}
\right)=v(\mathbf r, 0)(t+1)e^{-\zeta t}+v'(\mathbf r,0)te^{-\zeta t}+\int_0^t
{h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'} 

@@

	\label{eq:SecondOrderSynapticRespKernel} h(t) = At\exp{\left(-\zeta
t\right)} 

@@

where $\zeta=\tau^{-1}$, $\tau$ is the synaptic time constant.  $A$ determines
the maximal amplitude of the post-synaptic potentials.
Let $W(\mathbf r,t)$ be a spatio-temporal Wiener process. Then, for $\sigma \in
\mathbb R^{+}$, the non-local interactions $g\left( {\mathbf{r},t} \right)$
between cortical populations can be divided into a predominantly deterministic
part $g'\left( {\mathbf{r},t} \right) $ and  the random component $W(\mathbf
r,t)$,

@@

where $f(\cdot)$ is the firing rate function, $w(\cdot)$ is the spatial
connectivity kernel and $\Omega$ is the spatial domain representing a cortical
sheet or surface. Substituting \eqref{eq:SecondOrderActPotComponents} in
\eqref{eq:SecondOrderSpikesToPotential} we get

@@

It can be shown that the post-synaptic potential $v(\mathbf r, t)$ solves the
following initial value problem \parham{Andrew do you know how?}
